Layered Formal Verification of a TCP/IP Stack

### 2021-06-03 SecDev21/OryxStack.tex
fixing the second bug text

    \caption{Overview of the TCP/IP CycloneTCP stack in the OSI model. Green parts represent
             what have been translated in SPARK and in orange the underlying untrusted part written in C.}
    \label{Fig:TcpStack}
\end{figure}

>>>>

    \caption{Overview of the TCP/IP CycloneTCP stack (OSI model). Green parts have been translated into SPARK. They depend on orange parts written in C.}
    \label{Fig:TcpStack}
\end{figure}

### 2021-06-03 SecDev21/paper.tex
fixing the second bug text

\todo[inline]{KG: fixed up to here}
\subsubsection{Violation of the TCP protocol}

A subtle bug found thanks to SPARK is
reproduced in the snippet below.

@@

In this scenario, after the call to \spark{Tcp_Send}, when the error cases have
been filtered, the state of the socket can either be \sestab or \sclosew. After the call
at line~\ref{bugProg:tcpWaitForEvents}, the state of the socket can
still be \sestab or \sclosew, or it can be \sclosed if a \rst flag has been received when the mutex
was released.
At line~\ref{bugProg:tcpWaitForEventsPost}, a \fin segment is sent, which does
not modify the state of the socket. Finally, at line~\ref{bugProg:tcpChangeState}
we try to jump to the state \sfwone and thus the transitions $\text{\sclosew}
\rightarrow \text{\sfwone}$ and $\text{\sclosed} \rightarrow \text{\sfwone}$ are
possible.

But these transitions are not allowed by RFC 793, and thus not allowed by the
precondition of \spark{TCP_Change_State}. The problem is detected by SPARK and a fix
has been applied to the SPARK and C implementations of CycloneTCP.

>>>>

\subsubsection{Violation of the TCP protocol}

A subtle bug that violates the protocol's extracted specifications captured by SPARK is reproduced below:

@@

In the original implementation, after the call to \spark{Tcp_Send}, if no error occurs, the state of the socket can either be \sestab or \sclosew. After the call at line~\ref{bugProg:tcpWaitForEvents}, the state of the socket can still be \sestab or \sclosew, or \sclosed if a \rst flag has been received when the mutex was released. At line~\ref{bugProg:tcpWaitForEventsPost}, a \fin segment is sent, which does not modify the state of the socket. Finally, at line~\ref{bugProg:tcpChangeState} the code tries to move the connection to the \sfwone state, and thus, the transitions $\text{\sclosew} \rightarrow \text{\sfwone}$ and $\text{\sclosed} \rightarrow \text{\sfwone}$ are possible.

These transitions are not allowed by RFC 793, and thus, not allowed by the precondition of \spark{TCP_Change_State}, which models all the allowed transitions by the protocol's specification. The problem was both corrected in the C original implementation and the SPARK translated code.
